---
title: "Bærum sammendragsstatistikk"
author: "David Barton, Bart Immerzeel og Willeke A'Campo"
date: "13. november 2023"
#output:
#  html_document: default
#  pdf_document:  default
sansfont: Arial
params:
  region_str:
    label: "Kommune"
    value: Bærum
    input: select
    choices: ["Bærum", "Bodø", "Kristiansand", "Oslo"]
  region_var:
    label: "Kommune variabel"
    value: baerum
    input: select
    choices: ["baerum", "bodo", "kristiansand", "oslo"]
  data: 
    label: "Input dataset:"
    value: bærum_registrerte_traer.csv
    input: file
  agol:
    label: Bytraeratlas link (AGOL)
    tag: ['baerum', 'bodo', 'kristiansand', 'oslo']
    value: 8e112760eff34fd5b9176cefb7d31eb3
    input: select
    choices: ['8e112760eff34fd5b9176cefb7d31eb3', 
    '5191adc2c4b34658aea227c9853c6ebb', 
    '6e047c5432e64b3f9abb1592d7907ff6', 
    'oslo']
---

```{r "SETUP", echo=FALSE, results='hide',message=FALSE, warning=FALSE}
cwd <- getwd()
project_path <- cwd #dirname(cwd)


# INSTALL AND LOAD R PACKAGES
# ---------------------------
libs <- c("modules", "logger", "extrafont", "ggplot2", "arrow", "magrittr")

.installed_libs <- libs %in% rownames(
    installed.packages()
)


if (any(.installed_libs == FALSE)) {
    install.packages(libs[!.installed_libs])
}

invisible(lapply(
    libs, 
    library, 
    character.only = TRUE
))

logger::log_info(paste0("Project path: ", project_path))
logger::log_info(paste0("R packages loaded: ", paste(libs, collapse = ", ")))

# LOAD LOCAL MODULES
# ------------------
modules <- c(
    "config",
    "nina_theme",
    "nina_plot")

.load_module <- function(module_name) {
    modules::use(paste0(project_path, "/R/", module_name, ".R"))
}

invisible({
  modules_list <- lapply(modules, .load_module)
  names(modules_list) <- modules

  list2env(modules_list, envir = .GlobalEnv)
})

logger::log_info(paste0("Local modules loaded: ", paste(modules, collapse = ", ")))

# ENCODING 
# ------
Sys.setlocale("LC_CTYPE", "en_US.UTF-8")
logger::log_info(paste0("Encoding: ", Sys.getlocale("LC_CTYPE")))

# PARAMETERS
# ----------
#' Merge params from parameters.yaml with params from Rmd header
.update_params <- function(project_path) {
  yaml_parameters <- config$load_parameters(
      parameters_filepath = file.path(project_path, "/config/parameters.yaml")
  )
    
  parameters <- modifyList(params, yaml_parameters)

  assign("parameters", parameters, envir = .GlobalEnv)
  
  return(parameters)
}

parameters <- .update_params(project_path)
municipality = parameters$region_var
logger::log_info(paste0("Parameters configured for: ", municipality))
params_municipality = parameters[[municipality]]

# CATALOG
# ------
catalog <- config$load_catalog(
    catalog_filepath = file.path(project_path, "/config/catalog.yaml")
 )

catalog_municipality <- catalog[[municipality]]
```

```{r STYLE, echo=FALSE, results='hide',message=FALSE, warning=FALSE}
# STYLE
# -----
myfontPath <- catalog$fonts$filepath
myfontName <- "Arial"
config$set_font(font_path = myfontPath, font_name = myfontName, base_size = 12)


# NINA STYLE
# ----------
nina_grafpalett <- unname(
  NinaR::ninaColors( "purple", "orange", "dark blue", "light blue")
  )
nina_grafpalett_80 <- paste0(nina_grafpalett, "33")
nina_grafpalett_60 <-  paste0(nina_grafpalett, "66")
nina_grafpalett_40 <- paste0(nina_grafpalett, "99")
nina_grafpalett_20 <- paste0(nina_grafpalett, "CC")
nina_grafpalett <- c(nina_grafpalett, nina_grafpalett_20,
                     nina_grafpalett_40, nina_grafpalett_60, 
                     nina_grafpalett_80)
```

```{r "LOAD FUNCTIONS", echo=FALSE, results='hide',message=FALSE, warning=FALSE}
# FUNCTIONS
# ---------

# Function ------------------------------------------------------------------- #
#' Calculate summary statistics
#' @param data: data frame
#' @param x: variable name
#' @param classify_other: logical, if TRUE, classify all values with less than 3%
#' @return: a data frame with the summary statistics
#' @example
#' data <- data.frame(x = c("A", "B", "C", "C", "C"))
#' calc_summary(data, "x", classify_other = TRUE)
#' # A tibble: 3 x 3
#'  x         n  Perc
#' <chr> <int> <dbl>
#' 1 C         3  60
#' 2 A         1  20
#' 3 B         1  20

calc_summary <- function(data, x, classify_other = FALSE) {
    # drop "" and " " and "NA"
  data <- data %>%
    dplyr::filter(!!rlang::sym(x) != "" & 
                      !!rlang::sym(x) != " " & 
                      !is.na(!!rlang::sym(x)))

  # create summary dataframe df[x, n, perc]
  summary <- data %>%
    dplyr::count(!!rlang::sym(x)) %>%
    dplyr::mutate(Probability = n / sum(n),
                  Perc = round(Probability * 100, digits = 2)) %>%
    dplyr::arrange(desc(Perc))

  if (abs(sum(summary$Probability) - 1) > .Machine$double.eps^0.5) {
    stop("Error: Probability SUM != 1")
  }

  data_summary <- summary %>% dplyr::select(-Probability)

  if (classify_other) {
    data_summary <- summary %>%
      dplyr::filter(Perc >= 3) %>%
      dplyr::bind_rows(dplyr::summarise(dplyr::filter(summary, Perc < 3), 
                          # !! is used to unquote the variable name 
                          # var_name <- "my_var" unquote: my_var 
                          !!rlang::sym(x) := "Andre treslag*", 
                          n = sum(n), 
                          Perc = sum(Perc)))
    return(data_summary)
  }

  return(data_summary)
}

# Function ------------------------------------------------------------------- #
#' Calculate mean per class
#' @param df Data frame
#' @param category Category (tree species, DBH class, crown area class)
#' @param value_col Value column
#' @return Data frame with mean per class

.mean_per_class <- function(df, category, value_col, digits = 0) {
  df %>% 
    # remove NA 
    dplyr::filter(!is.na({{ value_col }})) %>%
    dplyr::group_by(!!sym(category)) %>% 
    dplyr::summarise(mean = round(mean(!!sym(value_col), na.rm = TRUE), digits)) %>% 
    dplyr::arrange(desc(mean))
}

# FUNCTION ------------------------------------------------------------------- #
#' Create a continuous  color palette
#' @param df Data frame
#' @param column Column name
#' @param colors Vector of three colors

create_color_palette <- function(df, 
                                 column,
                                 colors= c("#015c55","#7da37b", "#fae9a0")) {
  palette_func <- colorRampPalette(colors)
  n_classes <- length(unique(df[[column]]))
  palette <- palette_func(n_classes)

  return(palette)
}
```

```{r "LOAD DATA", echo=FALSE, results='hide',message=FALSE, warning=FALSE}
# LOAD DATA
# TODO check convert functions (they do not convert datatypes!)
# ---------

lookup_attr <- readxl::read_excel(
    path = catalog$lookup_attr$filepath,
    sheet = catalog$lookup_attr$sheet_1,
    col_names = TRUE,
    col_types = NULL,
    na = "",
    skip = 0
    )

df <- config$convert_excel_to_parquet(
    excel_file = catalog_municipality$bytraer$filepath,
    pqt_file = catalog_municipality$bytraer$filepath_parquet,
    lut_df = lookup_attr
    )

df_2 <- config$convert_csv_to_parquet(
    csv_file = catalog_municipality$trekroner$filepath,
    pqt_file = catalog_municipality$trekroner$filepath_parquet,
    sep = ";",
    lut_df = lookup_attr
    )


message <- paste(
    "Data loaded.",
    paste0("DF dimensions: ", 
           paste(dim(df), collapse = " x ")),
    paste0("DF2 dimensions: ", 
           paste(dim(df_2), collapse = " x ")),
    paste0("Lookup name-alias: ", 
           lookup_attr[1,1]),
    sep = "\n"
  )
logger::log_info(message)
```

```{r "DATA MANIPULATION", echo=FALSE, results='hide',message=FALSE, warning=FALSE}
# CLEAN DATA
# ----------

# 0. Set DTYPES
# -------------
# set dtype pollution_zone to int
df$pollution_zone <- as.integer(df$pollution_zone)

# ITree Eco is not run for Baerum 
pos_cols <- c("height_total_tree", "dbh", "crown_area")

df[pos_cols] <- lapply(df[pos_cols], function(x) round(as.numeric(x), 2))
df[pos_cols][df[pos_cols] < 0] <- NA

# 1. Create <ID>
# --------------
# # tree_id: itree_3006 ID_ 3006
df$ID <- as.numeric(gsub("itree_", "", df$tree_id))
df <- df[order(df$ID),]

# 2. Create <class>
# -----------------
df$class <- "Alle traer"

# 3. Tre species column 
# ---------------------------
# if Hule eiker > eik
df$norwegian_name <- ifelse(df$norwegian_name == "Hule eiker", "eik", df$norwegian_name)

# if skjøttetTre or gammeltTre or hultTre, andreTreslag set to NaN
df$norwegian_name <- ifelse(df$norwegian_name == "skjøttetTre", NA, df$norwegian_name)
df$norwegian_name <- ifelse(df$norwegian_name == "gammeltTre", NA, df$norwegian_name)
df$norwegian_name <- ifelse(df$norwegian_name == "hultTre", NA, df$norwegian_name)
df$norwegian_name <- ifelse(df$norwegian_name == "andreTreslag", NA, df$norwegian_name)

# print unique nowegian_name
print(paste0("Unique tree species: ", toString(unique(df$norwegian_name))))

logger::log_info(paste0("Data cleaned \n", toString(df[1,3:6])))

# FILTER DATA
# -----------

# 1. COUNT DBH, NORWEGIAN_NAM
# ---------------------------
# number of trees with DBH 
n_trees_dbh <- df %>% 
  dplyr::filter(!is.na(dbh)) %>% 
  dplyr::count(dbh) %>% 
  dplyr::summarise(n = sum(n))

n_trees_species <- df %>% 
  dplyr::filter(!is.na(norwegian_name)) %>% 
  dplyr::count(norwegian_name) %>% 
  dplyr::summarise(n = sum(n))

# 2. Zone 1 trees
# ---------------
df_z1 <- df[df$pollution_zone == 1,]
df_z1$class <- "Sone 1"

# 3. Zone 2 trees
# ---------------
df_z2 <- df[df$pollution_zone == 2,]
df_z2$class <- "Sone 2"

# 4. Zone 3 trees
# ---------------
df_z3 <- df[df$pollution_zone == 3,]

logger::log_info("Data grouped into separate dataframes:
                 \ndf (Alle traer), df_itree (i-Tree), 
                 df_z1 (Sone 1), df_z2 (Sone 2), df_z3 (Sone 3)")

df_list = list(df = df, 
               df_z1 = df_z1, 
               df_z2 = df_z2,
               df_z3 = df_z3)

# MERGE DATA
# ----------
data_merged <- rbind(df, df_z1, df_z2, df_z3)
dat <- tibble::as_tibble(data_merged)

logger::log_info("Data merged into one tibble: dat")
```

### Registrerte trær i `r params$region_str`'s byggesonen

Dette dokumentet viser sammendragstatistikken for registrerte trær innenfor `r params$region_str`'s byggesone. Tilknyttede kartprodukter er synlige i bytræratlaset: [Bytræratlas `r params$region_str`](https://experience.arcgis.com/experience/%60r%20params$agol%60/)

| Gruppe                 | Beskrivelse                                            | Antall          |
|------------------|-----------------------------------|------------------|
| Totalt registerte trær | Totalt antall registrerte trær i `r params$region_str` | `r nrow(df)`    |
| Treslag                | Trær med treslag                                       | `r n_trees_species`            |
| Stammediameter         | Trær med stammediameter                                | `r n_trees_dbh`             |
| i-Tree Eco             | Totalt antall trær brukt i i-Tree Eco                  | 0               |
| Sone 1                 | Trær i forurensnings sone 1 (grønn)                    | `r nrow(df_z1)` |
| Sone 2                 | Trær i forurensnings sone 2 (gull)                     | `r nrow(df_z2)` |
| Sone 3                 | Trær i forurensnings sone 3 (rød)                      | `r nrow(df_z3)` |
| Totalt trebestand      | Totalt antall trær i bestanden (basert på laserdata)   | `r nrow(df_2)`  |

<br>

### Treslagsfordeling i `r params$region_str`

Treslag som har en sannsynlighet på mindre enn 3% er klassifisert som "Andre treslag".

```{r "SPECIES_PROBABILITY", echo=FALSE, results='hide',message=FALSE, warning=FALSE, fig.height=15, fig.width=14}
# CALCULATE SPECIES PROBABILITY
# -----------------------------
# 1. Calc probability for each tree species per group
# Groups: "Alle Traer", "i-Tree", "Sone 1" and "Sone 2"
summary_list <- lapply(
    X = df_list, 
    FUN = function(df) calc_summary(
        df, 
        "norwegian_name", 
        classify_other = params_municipality$norwegian_name$classify_other))

# 2. Get species with probability < 3%
other_trees <- df %>%
    dplyr::filter(!!rlang::sym("norwegian_name") != "" & 
                  !!rlang::sym("norwegian_name") != " " & 
                  !is.na(!!rlang::sym("norwegian_name"))) %>%
    dplyr::count(norwegian_name) %>%
    dplyr::mutate(Probability = n / sum(n)) %>%
    dplyr::filter(Probability < 0.03) %>%
    dplyr::pull(norwegian_name)

# PLOT SPECIES PROBABILITY
# ------------------------
plot_all <- nina_plot$probability_plot(
    data = summary_list$df,
    Var1 = "norwegian_name",
    title = paste0("Trær i ", params$region_str, "'s byggesone"),
    ymax = params_municipality$norwegian_name$ymax  
    ) +
    theme_minimal(base_family = "Arial", base_size = 16) +
    nina_theme$theme_large()

plot_z1 <- nina_plot$probability_plot(
    data = summary_list$df_z1,
    Var1 = "norwegian_name",
    title = "Trær i forurensningssone 1",
    ymax = params_municipality$norwegian_name$ymax
    ) +
    theme_minimal(base_family = "Arial", base_size = 16) +
    nina_theme$theme_large()

plot_z2 <- nina_plot$probability_plot(
    data = summary_list$df_z2,
    Var1 = "norwegian_name",
    title = "Trær i forurensningssone 2", 
    ymax = params_municipality$norwegian_name$ymax
    ) +
    theme_minimal(base_family = "Arial", base_size = 16) +
    nina_theme$theme_large()

plot_z3 <- nina_plot$probability_plot(
    data = summary_list$df_z2,
    Var1 = "norwegian_name",
    title = "Trær i forurensningssone 3",
    ymax = params_municipality$norwegian_name$ymax
    ) +
    theme_minimal(base_family = "Arial", base_size = 16) +
    nina_theme$theme_large()

plots <- ggpubr::ggarrange(
    plot_all, 
    plot_z1, 
    plot_z2,
    plot_z3,
    ncol = 2, nrow = 2, align = "v"
)

print(plots)
```

<font size="0.8" face="Arial">\*Treslag klassifisert som 'Andre treslag' er: `r other_trees` </font>

<br>

### Tre egenskaper

------------------------------------------------------------------------

```{r "TREE_ATTRIBUTES", echo=FALSE, results='hide',message=FALSE, warning=FALSE, fig.height=4, fig.width = 14}
# PLOT HISTOGRAM: CROWN AREA, HEIGTH, DBH
# ---------------------------------------

# 1. Prepare plot parameters
# --------------------------
prepare_plot <- function(df_list, variable_name, params_municipality) {
  
    # Calc histogram parameters
    hist_params_list <- lapply(
        X = df_list, 
        FUN = function(df) nina_plot$calculate_bins(
            df = df,
            column_name = variable_name, 
            bin_method = "fd"))

    # Set plot params for each data frame
    params_histogram <- list(
        hist_params_list$df, 
        hist_params_list$df_z1, 
        hist_params_list$df_z2, 
        hist_params_list$df_z3)

    # Investigate params and set manual values to improve plot
    new_bin_max <- params_municipality[[variable_name]]$bin_max
    new_bin_step <- params_municipality[[variable_name]]$bin_step
    new_ymax <- params_municipality[[variable_name]]$ymax
    titles <- params_municipality[[variable_name]]$title
  
    for (i in seq_along(params_histogram)) {
        params_histogram[[i]]$bin_max <- new_bin_max[i]
        params_histogram[[i]]$bin_step <- new_bin_step[i]
        params_histogram[[i]]$ymax <- new_ymax[i]
        params_histogram[[i]]$title <- titles[i]
    }
    return(params_histogram)
}

# 2. Plot histograms
# ------------------
plot <- function(params_histogram, variable_name, main_title) {
    
    # loop through each data frame
    for (i in seq_along(params_histogram)) {
        logger::log_info("Plotting histogram for ", main_title, " - ", params_histogram[[i]]$title)
        #logger::log_info("Plotting params:")
        #logger::log_info("bin_max: ", params_histogram[[i]]$bin_max)
        #logger::log_info("bin_step: ", params_histogram[[i]]$bin_step)
        #logger::log_info("ymax: ", params_histogram[[i]]$ymax)
        
        x_axis_title <- params_municipality[[variable_name]]$x_axis_title

        # Create the subplots
        plots <- lapply(seq_along(params_histogram), function(i) {
            nina_plot$histogram(
                data = params_histogram[[i]]$df_capped, 
                value_col = variable_name, 
                bin_max = params_histogram[[i]]$bin_max,
                bin_step = params_histogram[[i]]$bin_step, 
                ymax = params_histogram[[i]]$ymax,
                title = params_histogram[[i]]$title,
                axis_title_x = x_axis_title
            ) + 
                theme_minimal(base_family = "Roboto", base_size = 12) +
                nina_theme$theme_medium()
        })
        
        # Combine the subplots
        sub_plot <- ggpubr::ggarrange(
            plots[[1]], 
            plots[[2]], 
            plots[[3]], 
            plots[[4]],
            ncol = 4, nrow = 1, align = "v"
        )
        
        # Create titles
        # Create the labels
        title <- cowplot::ggdraw() + cowplot::draw_label(main_title, 
                                       fontfamily = "Arial", 
                                       x = 0.45, y = 0.2, 
                                       hjust = 0, vjust = 0, 
                                       size = 14, color = "#000000",
                                       fontface = "bold")
        # Combine the labels and the plots
        main_plot <- cowplot::plot_grid(title, sub_plot, ncol = 1, rel_heights = c(0.2, 1))
    }

  return(main_plot)
}

# prepare plot parameters
params_histogram_CA <- prepare_plot(df_list, "crown_area", params_municipality)
params_histogram_H <- prepare_plot(df_list, "height_total_tree", params_municipality)
#params_histogram_DBH <- prepare_plot(df_list, "dbh", params_municipality)


# plot
plot_CA <- plot(params_histogram_CA, "crown_area", "Kroneareal")
plot_H <- plot(params_histogram_H, "height_total_tree", "Høyde")
#plot_DBH <- plot(params_histogram_DBH, "dbh", "Stammediameter") # NOT ENDOUGH INFO

print(plot_CA)
print(plot_H)
#print(plot_DBH)

# write plots to file
# code here
```

```{r "summary_table", echo=FALSE, results='hide',message=FALSE, warning=FALSE}
# Assuming df_itree is your data frame and it has columns for tree species, crown area, height, and stem diameter
library(dplyr)
summary_df <- df %>%
  group_by(Treslag = norwegian_name) %>%
  summarise(
    `Antall Trær` = n(),
    `Kroneareal (m2)` = round(mean(crown_area, na.rm = TRUE),2),
    `Høyde (m)` = round(mean(height_total_tree, na.rm = TRUE),2),
    `Stammediameter (cm)` = round(mean(dbh, na.rm = TRUE),2)
  )
# replace 'NaN' with " "


# Print the summary_df in a format suitable for markdown
table_md <- knitr::kable(summary_df, format = "markdown")
writeLines(table_md)

# export to csv)
path <- file.path(project_path, "reports", paste0(params$region_var, "_summary_table_structrual_values.csv"))
print(path)
write.csv(summary_df, file = path)
```

```{r "knitt", echo=FALSE, results='hide',message=FALSE, warning=FALSE}  
knitr::knit('baerum_summary_statistics.Rmd', 'baerum_summary_stat.md')
```

<br>

*Dette dokumentet er en del av prosjektet:*

**TREKRONER Prosjektet** \| Trærs betydning for klimatilpasning, karbonbinding, økosystemtjenester og biologisk mangfold.
